---
params:
  set_title: ""
  yearofsurvey: 
title: "`r params$set_title`"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
---

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide'}
#######################################
# PROJECT: More Community, Less Confinement (2024)
# PURPOSE: Impute values, national estimates, state estimates and costs
# AUTHOR: JSM
#######################################

###################################################
#OVERALL SETUP
#load necessary packages
library(openxlsx)
library(dplyr)
library(readr)
library(reshape)
library(readxl)
library(tidyverse)
library(data.table)
library(formattable)
library(scales)
library(mice)
library(VIM)
library(finalfit)
library(janitor)
library(Hmisc)
library(eeptools)
library(csgjcr)
library(scales)
library(xtable)

#print to word
library(Gmisc)

#for binding blank dataframes
cbind.fill <- function(...){
  nm <- list(...) 
  nm <- lapply(nm, as.matrix)
  n <- max(sapply(nm, nrow)) 
  do.call(cbind, 
          lapply(nm, function (x) rbind(x, matrix( , n-nrow(x), ncol(x))))
          ) 
  }

##CHANGE INPUT BELOW!!!#####################################
#
MCLCsurveyyear <- params$yearofsurvey
#
##CHANGE INPUT ABOVE!!!#####################################

#years of data
year     <- as.character(2018:MCLCsurveyyear)
refyear  <- as.character(2019:MCLCsurveyyear)

#for labeling tables in output
tablevals.pretty <- c('Overall Admissions',
                     'Admissions for Probation Violations',
                     'Admissions for Parole Violations',
                     'Total Admissions for Technical Violations',
                     'Admissions for Technical Probation Violations',
                     'Admissions for Technical Parole Violations',
                     'Total Admissions for New Crime Violations',
                     'Admissions for New Crime Probation Violations',
                     'Admissions for New Crime Parole Violations',
                     
                     'Overall Population',
                     'Population of Probation Violators',
                     'Population of Parole Violators',
                     'Total Population of Technical Violators',
                     'Population of Technical Probation Violators',
                     'Population of Technical Parole Violators',
                     'Total Population of New Crime Violators',
                     'Population of New Crime Probation Violators',
                     'Population of New Crime Parole Violators'
                     )

##import data
source("import.R")

################################################################################
# IMPUTATION
# MICE
# National Estimates w/95% CI's
################################################################################
yearnum <- 1:length(year)
refnum  <- 1:(max(yearnum)-1)

#calculate imputed values (multiple imputation)
source("MImp1.R")

######
#produce confidence intervals
#post multiple imputation t-test
#grab response columns for creating estimated counts/CIs
######

source("MImp2.R")

##############################################################
################## CREATE A TABLE HERE #######################
table2018 <- cbind(nat.2018all,natCI.2018all) %>%
  mutate(across(where(is.numeric),formattable::comma,0),
         id     = numvar,
         ' '    = '95% Confidence Interval',
         '2018' = paste0('(',`V1`,', ',`V2`,')')
         )

for (i in 2019:MCLCsurveyyear) {
  temp1 <- get(paste0("nat.",i,"all"))
  temp2 <- get(paste0("natCI.",i,"all"))
  test  <- cbind(temp1,temp2) %>%
    mutate(across(where(is.numeric),formattable::comma,0),
           !!as.name(i) := paste0('(',`V1`,', ',`V2`,')'))
  assign(paste0("table",i),test,envir = .GlobalEnv)
}

#estimates
tab.full1 <- data.table()
for (i in year) {
  temp <- get(paste0("table",i))$Estimate
  tab.full1  <- data.table(cbind(tab.full1,temp))
  setnames(tab.full1, "temp", i)
  tab.full1 <- tab.full1[, lapply(.SD, as.character)]
}

#confidence intervals
CIS <- data.table(' '  = table2018$` `,
                  'id' = table2018$id)
for (i in year) {
  temp <- get(paste0("table",i))[[i]]
  CIS  <- cbind(CIS,temp)
  setnames(CIS, "temp", i)
}

#combine
tab.full <- tab.full1 %>%
  mutate(id = numvar,
         ' ' = 'Estimate') %>%
  relocate(any_of(c('id',' '))) %>%
  rbind(
    data.table(' ' = tablevals.pretty,
               id  = numvar,
               id2 = 0),
    CIS,
    fill = TRUE) %>%
  mutate(id     = as.numeric(id)) %>%
  arrange(id,id2) %>%
  select(-c(id,id2))

#############################################################################
###############CALCULATING STATE ESTIMATES
#############################################################################

source("MImp3.R")

#############################
###COSTS#####################

source("Costs.R")

```

## National Estimates

These estimates incorporate BJS total population and admissions data in calculations across all years.

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}

print(xtable(tab.full[1:((max(numvar)/2)*3),],
                           align = c("llcccccc")),
      include.rownames = FALSE,
      type             = "html")

```

<br>
<br>

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
print(xtable(tab.full[(((max(numvar)/2)*3)+1):(((max(numvar)/2)*3)*2),],
                           align = c("llcccccc")),
      include.rownames = FALSE,
      type             = "html")
```

## `r as.numeric(last(refyear))` Costs

The total annual cost for incarcerating **probation violators** in `r as.numeric(last(refyear))` is `r dollar_format()(total.pr.vp)`. The total annual cost for incarcerating **parole violators** in `r as.numeric(last(refyear))` is `r dollar_format()(total.pa.vp)`.

<br>

\noindent The total annual cost for incarcerating just those with **technical violations** in `r as.numeric(last(refyear))` is `r dollar_format()(total.tvp)`.

<br>

\noindent The total annual cost for incarcerating just those with **new offense violations** in `r as.numeric(last(refyear))` is `r dollar_format()(total.novp)`.

<br>

### `r as.numeric(last(refyear))` Costs by State

Below is a table of these costs by state (no imputation). Note that missing cells indicate that a state did not submit `r as.numeric(last(refyear))` population data:

<br>

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
print(xtable(cost.final[c("State","cost.pr.state","cost.pa.state","cost.tpr.state","cost.tpa.state")] %>%
                              mutate(
                                `Overall Probation Violator Cost`   = case_when(cost.pr.state >=0 ~ dollar_format()(cost.pr.state),
                                                                                  TRUE ~ NA_character_),
                                `Overall Parole Violator Cost`      = case_when(cost.pa.state >=0 ~ dollar_format()(cost.pa.state),
                                                                                  TRUE ~ NA_character_),
                                `Technical Probation Violator Cost` = case_when(cost.tpr.state >=0 ~ dollar_format()(cost.tpr.state),
                                                                                  TRUE ~ NA_character_),
                                `Technical Parole Violator Cost`    = case_when(cost.tpa.state >=0 ~ dollar_format()(cost.tpa.state),
                                                                                  TRUE ~ NA_character_)
                                     ) %>%
                              select(-c(cost.pr.state,cost.pa.state,cost.tpr.state,cost.tpa.state)),
                           align = c("llrrrr")),
      include.rownames = FALSE,
      type             = "html")
```

## `r as.numeric(last(refyear)) - 1` Costs

The total annual cost for incarcerating **probation violators** in `r as.numeric(last(refyear)) - 1` is `r dollar_format()(total.pr.vp22)`. The total annual cost for incarcerating **parole violators** in `r as.numeric(last(refyear)) - 1` is `r dollar_format()(total.pa.vp22)`.

<br>

\noindent The total annual cost for incarcerating just those with **technical violations** in `r as.numeric(last(refyear)) - 1` is `r dollar_format()(total.tvp22)`.

<br>

\noindent The total annual cost for incarcerating just those with **new offense violations** in `r as.numeric(last(refyear)) - 1` is `r dollar_format()(total.novp22)`.

<br>

### `r as.numeric(last(refyear)) - 1` Costs by State

Below is a table of these costs by state (no imputation). Note that missing cells indicate that a state did not submit `r as.numeric(last(refyear)) - 1` population data:

<br>

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
print(xtable(cost.final[c("State","cost.pr.state22","cost.pa.state22","cost.tpr.state22","cost.tpa.state22")] %>%
                              mutate(
                                `Overall Probation Violator Cost`   = case_when(cost.pr.state22 >=0 ~ dollar_format()(cost.pr.state22),
                                                                                  TRUE ~ NA_character_),
                                `Overall Parole Violator Cost`      = case_when(cost.pa.state22 >=0 ~ dollar_format()(cost.pa.state22),
                                                                                  TRUE ~ NA_character_),
                                `Technical Probation Violator Cost` = case_when(cost.tpr.state22 >=0 ~ dollar_format()(cost.tpr.state22),
                                                                                  TRUE ~ NA_character_),
                                `Technical Parole Violator Cost`    = case_when(cost.tpa.state22 >=0 ~ dollar_format()(cost.tpa.state22),
                                                                                  TRUE ~ NA_character_)
                                     ) %>%
                              select(-c(cost.pr.state22,cost.pa.state22,cost.tpr.state22,cost.tpa.state22)),
                           align = c("llrrrr")),
      include.rownames = FALSE,
      type             = "html")
```
